---
title: 'STAT847 Lab 5'
author: "Anh Pham - ewa589"
date: 'October 5, 2023'
output:
  html_document:
    theme: united
    toc: yes
    toc_float: no
    number_sections: no
    highlight: tango
    fig_width: 10
    fig_height: 8
  word_document:
    toc: yes
  pdf_document:
    toc: yes
    fig_height: 6
    fig_width: 10
    number_sections: no
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
1) Load in the "agpop.csv" file to an object called 'ag.data'
```{r}
ag.data <- read.csv('D:agpop-1.csv')
head(ag.data)
```
2) Use the `summary()` function to print out one summary of `acres92` and `acres87`
```{r}
summary(ag.data[c('acres92','acres87')])
```
3) Print out a histogram of `acres92` using the base R `hist()` function.  Change the title of the histogram to "Histogram of Total Acres in 1992". Change the x-axis label to "Acres in 1992"
```{r}
hist(ag.data$acres92, main = "Histogram of Total Acres in 1992", xlab = "Acres in 1992")
```
4) Use the `sample()` function to create sampling indexes that split the 'ag.data' into 80% for training and 20% for testing. Then use these sample indexes to split the 'ag.data' into training and testing datasets called "train.data" and "test.data".
```{r}
set.seed(1)
dt= sort(sample(nrow(ag.data), nrow(ag.data)*.8))
train.data <- ag.data[dt,]
test.data <- ag.data[-dt,]
dim(train.data)
dim(test.data)
```