---
title: 'STAT847 Lab 4'
author: "Anh Pham - ewa589"
date: 'September 28, 2023'
output:
  pdf_document:
    toc: yes
    fig_height: 6
    fig_width: 10
    number_sections: no
  word_document:
    toc: yes
  html_document:
    theme: united
    toc: yes
    toc_float: no
    number_sections: no
    highlight: tango
    fig_width: 10
    fig_height: 8
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1
Load the `mtcars.csv` file into an object names "mtcars.data". You can use `?mtcars` to get more information on the data set
```{r}
mtcars.data <-read.csv('D:mtcars.csv')
head(mtcars.data)
?mtcars
```
# Question 2
Fit a logistic regression model to predict `am`(transmission) using `mpg`, `disp`, `hp`, and `wt`. 
```{r}
myfit <- glm(am ~ mpg + disp + hp + wt, data = mtcars.data,
family = "binomial")
summary(myfit)
```
Print out the model coefficients
```{r}
coef(myfit)
```
# Question 3
Use the `predict()` function to get the fitted probabilities. Then convert the probabilities to 1 or 0 using the threshold of 0.5.  For example, if the fitted probability is greater than 0.5, it would equate to a 1 values, if the fitted probability is less than or equal to 0.5 it equates to a 0 value.

First, we get the fitted probability
```{r}
glm.probs=predict(myfit, type="response")
head(glm.probs)
```
Now we convert the probilities to 1 or 0 using the threshold of 0.5
```{r}
glm.pred<- ifelse(glm.probs > 0.5,1,0)
head(glm.pred)
```
# Question 4
Create a confusion matrix to assess the accuracy of the fitted logistic model

Create a confusion matrix
```{r}
table(glm.pred, mtcars.data$am)
```
True classification rate
```{r}
mean(glm.pred==mtcars.data$am)
```
mis-classification rate
```{r}
mean(glm.pred!=mtcars.data$am)
```